---
title: ""
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(viridis)
library(leaflet)
library(dplyr)
library(ggplot2)
library(ggmap)
library(sf)
library(mapview)
library(rgdal)
library(htmltools)
library(plotly)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6, 
  out.width = "90%"
)


options(
  ggplot2.countinuous.colour = "viridis",
  ggplot2.countinuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

## reading data
ufo_data = read_csv(file = "./data/tidied_data_final.csv")


ufo =
ufo_data %>%  
separate( date, into = c("month","day","year"), sep = "/")
```

# Data Description

In this dataset, it contains time and location data about UFO sighting across `r length(unique(pull(ufo, country)))` countries, from the year of `r min(pull(ufo, year))` to the year of `r max(pull(ufo, year))`. They are ` r nrow(ufo)` observations and `r ncol(ufo)` variables in the dataset. 

The variables in the dataset are:

* `year`: the year of UFO sighting
* `month`: the month of UFO sighting
* `day`: the day of UFO sighting
* `time`: the extra time of UFO sighting
* `city_description`: the extra description that came along with the city of UFO sighting
* `city`: the city if UFO sighting
* `state`: the state of UFO sighting
* `ufo_shape`: the shape of the sighted UFO
* `encounter_length`: the encounter time length of sighting UFO
* `described_encounter_length`: the description from the people who reported the UFO encounter
* `description`: the description of the UFO from the people who reported the UFO encounter
* `date_documented`: the date of documentation of the UFO encounter
* `latitude`: the latitude of UFO encounter
* `longtitude`: the longtitude of UFO encounter
* `country`: the country of UFO sighting


# UFO Data Overview. 

```{r}
## A bargraph showing the total number of UFO observed among 7 countries.
bar = 
ufo %>%
  group_by(country) %>%
  count() %>%
  ungroup()%>%
 mutate(precentage = n/sum(pull(., n)),
      country = factor(country),
      country = forcats::fct_reorder(country, n)) %>%
ggplot(aes(x = country, y = precentage , fill = country, label =  scales::percent(precentage))) + 
    geom_col(position = 'dodge') + 
    geom_text(position = position_dodge(width = .9),  
              vjust = -0.5,   size = 3) + 
    labs(title = "Bargraph of the precentage of UFO Observed by Country",
         x = "Country",y = "Precentage") 
ggplotly(bar)


```

According to the graph above, the majority, 92.4%, of the UFO were observed in the United States of America and smaller portion of data came from Canada and the United Kingdom. We can almost ignore data from Australia, Ireland, Germany, Isle of Man and Puerto Rico. For statisticaly significant analysis, we would like to include enough data to perform tests and we will have solid results, hopefully.


```{r}
## line plot showing the number of UFO observed across time among countries
line_plot = ufo %>%
group_by(year) %>%
count(country) %>%
ungroup() %>%
mutate(year = as.numeric(year)) %>%
  ggplot(aes(x = year,y = n,group = country)) +
  geom_line(aes(color = country)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
title = " Number of UFO Across Time in different Countries ",
    x = "Year",
    y = "Number Of UFO Observed"
  )

ggplotly(line_plot)

```

This line plot provides a trend showing the number of UFO observed in each country across time. According to the plot, it is obvious that USA (indicated by the yellow line) has the most number of data. We can see that for the United States, the earliest data happens before 1925. The amount of UFO sighting begins to increase sometime after 1950. Most of UFO sighting reports were taken place around 1990 and after. Similar trend was also observed in Canada (indicated by the violet line), Puerto Rico (indicated by the green line) and Australia (indicated by dark purple line), but with far less amount of data recorded. 



# Descriptive Statistics in Encountered Length

```{r}

ufo %>% 
  group_by(country) %>% 
  # create columns for count, mean, median, sd, Q1, Q3, IQR for each country
  summarize(count = n(),
            mean = mean(encounter_length),
            median = median(encounter_length),
            sd = sd(encounter_length),
            Q1 = quantile(encounter_length, 0.25),
            Q3 = quantile(encounter_length, 0.75),
            IQR = IQR(encounter_length)) %>% 
  # arrange according to mean for each group
  arrange(desc(mean)) %>% 
  # knit a table
  knitr::kable(digits = 2, 
               caption = "The mean, median, sd, Q1, Q3 and IQR are for the variable length time (in seconds). ")

```

With relative small number of observations in Ireland, Germany, Isle of Man, Australia and Puerto Rico, the rest of analysis will mostly focus on the USA. The United Kingdom has the largest mean, 66512.89 seconds, and largest standard deviation, 2322650.18, and the smallest mean is Isle of Man, 52.50 seconds. According to the table above, since the standard deviations of the United States and Canada are way too large for a “readable” display, also we have decided above that we are mainly focusing on these two countries for analysis, extreme values will be omitted for the later density plot. 


```{r}
# create dataframe for the plot (which got rid of outliers for each country)
filtered_ufo = ufo %>% 
  group_by(country) %>% 
  # add two columns according to its country: Q3, IQR
  mutate(
    Q3 = quantile(encounter_length, 0.75),
    IQR = IQR(encounter_length)
  ) %>% 
  # ungroup
  ungroup() %>% 
  # filter out outliers
  filter(encounter_length < Q3 + 1.5 * IQR)

# make a density plot
density_plot = 
filtered_ufo %>%
  # make x-aixs = encounter_length, fill the color by country
  ggplot(aes(x = encounter_length, fill = country)) +
  # set the transparency to 0.4 and set several customerized parameter
  geom_density(alpha = 0.4, adjust = 0.5, color = "cornsilk4") +
  # to make the fill colors a bit prettier
  scale_fill_viridis(option = "plasma", discrete = TRUE) +
  labs(title = "Density Plot of Length Time for Each Country",
       x = "Encounter Length Time")
ggplotly(density_plot)


```

In the plot, United States showed the most amount of data and spreaded out a lot with some spikes of encounter length. It is also important to mention that a lot of encounter happened within 500 seconds. In the United States, there seemed to be a pattern of sighting time (due to the spikes). Hence,a categorical analysis will be conducted later in the project. Meanwhile, a similar pattern was observed in Canada yet "milder" as they both shared similar spikes and distribution. Hence, questions like whether there is a particular pattern behind the UFO sighting might be asked. For other countries that were displayed in the plot above, we also observed similar pattern to some degree.



# Lag in Documentation:

In the dataset, we noticed that for a lot of the data, there are lags in time regards to their sighting time and the documentation time. Lags may indicate inaccurate information in the dataset, since people's memory tends to become unreliable after time. Therefore, here is a plot that shows the distribution of lag time in years for each sighting by country.

```{r}
# tidy data
lag_data = ufo %>% 
  # separate the date_documented into month, day and year
  separate(date_documented, into = c("doc_month","doc_day","doc_year"), sep = "/") %>% 
  # add a variable called lag
  mutate(lag = as.numeric(doc_year) - as.numeric(year))
  
# violin plot for the lag
lag_plot = 
lag_data %>% 
  ggplot(aes(x = country, y = lag)) +
  geom_violin(aes(fill = country), color = "cornsilk4", alpha = 0.4) +
  stat_summary(fun.y = median, geom = "point", color = "darksalmon", size = 3) +
  labs(title = "Violin Plot for the Lag Time in Actual Sighting Date and Documentation Date")

ggplotly(lag_plot)


```

Based on the plot, we can see that most of the data are documented relatively recent to their sighting, which is a good thing, because the longer the lag, the larger the possibility of wrongly documentation. However, there is oulier in most of countries, they are not common (only takes a very small part in the dataset). Hence, we consider that the most of dataset are reliable.

