---
title: "UFO in USA"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
try(source("./dependencies.R"), silent = TRUE)
set.seed(1)
library(tidyverse)
library(leaflet)
library(dplyr)
library(ggplot2)
library(ggmap)
library(sf)
library(mapview)
library(rgdal)
library(htmltools)
library(viridis)
library(ggridges)
library(patchwork)
library(rvest)
library(modelr)
library(mgcv)


knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6, 
  out.width = "90%"
)


options(
  ggplot2.countinuous.colour = "viridis",
  ggplot2.countinuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
ufo_data = readr::read_csv("./data/tidied_data_final.csv")
```

# Visualization of UFO in USA

The goal of this part is to explore whether there exist any association between the shape of UFOs and their appearance in the U.S, as well as to detect if there is any potential correlation between shape of UFOs and their encounter length time.



# The Distribution of UFO Shape in the U.S


<strong> Boxplot showing the distribution of the UFO Shape Observed in US </strong>.

```{r boxplot_shape_distribution}
### Boxplot showing the distribution of the UFO Shape Observed in US

ufo =
ufo_data %>%  
#separate(date_time, into = c( "date","time"), sep = " " ) %>%
separate( date, into = c("month","day","year"), sep = "/") %>% 
  filter(country == "USA",year >= 1950) %>%
  group_by(state) %>%
  count(ufo_shape) %>%
  arrange(desc(n)) %>% 
  ungroup %>% 
  mutate(
  ufo_shape = factor(ufo_shape),
  ufo_shape = forcats::fct_reorder(ufo_shape, n)) %>%
  ggplot(aes(x = ufo_shape, y = n, group = ufo_shape)) +
  geom_boxplot(aes(color =ufo_shape)) +
  theme(axis.text.x = element_text(angle = 300)) +
  labs(
    title = "Boxplot Showing Distribution of the UFO Shape Observed in the U.S.",
    x = "UFO Shape",
    y = "Number"
  ) +
  theme(plot.title = element_text(hjust = 0.5))

ufo

```


According to this plot, we can see that the light UFO shape was sighted most often, and the changed shape was the least often according to their medium value. Some of the most common sighted UFO shapes are light, trianle, circle, unknown, other, and fireball, disk, and sphere. Some of the least common sighted UFO shapes are changed crescent, delta, flare, hezagon, pyramid and round. 


After seeing the distribution of shape and UFO sighting frequency, we are trying to explore whether the frequency of UFO shape changes by day and night. In general, the UFOs are more likely to be observed during night and the shape of light is the most frequent shape during night as indicated by the plot below.


<strong> Times of Appearance for Each Shape of UFO (by day and night) </strong>

```{r day_night_shape}
ufo_day_night_shape =
ufo_data %>%  
separate(date, into = c("month","day","year"), sep = "/") %>%
separate(time, into = c("hour","minute"), sep = ":") %>% 
mutate(hour=recode(hour, "00"="24")) %>% 
filter(country == "USA") %>% 
mutate(year_1 = case_when(year >= 1950 ~ "true",
                          TRUE ~"false")) %>% 
filter(year_1 == "true") %>% 
select(-year_1) %>% 
mutate(hour = as.numeric(hour)) %>% 
mutate(night_day = case_when(18 <= hour & hour <= 24~ "night",
                             1 <= hour & hour <= 6 ~ "night",
                             TRUE ~ "day")) %>% 
group_by(ufo_shape, night_day) %>% 
summarize(count_shape = n()) %>% 
  ggplot(aes(x = ufo_shape, y = count_shape, group = night_day)) +
  geom_col(aes(fill = night_day), alpha = .75, position = position_dodge(1)) +
  scale_fill_manual(
    values = c("lightblue","lightpink"), 
    labels = c("day","night")
    ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Times of Appearance for Each Shape of UFO (by day and night)", 
       y = "counts",
       x = "UFO Shape") +
  theme(plot.title = element_text(hjust = 0.5))
  
  ufo_day_night_shape
```


The spagetti plot below confirms that UFOs are more likely to appear during night, since the frequency of appearance dramatically after 6pm. 


<strong> Numbers of appearance of UFO in each hour by month </strong>
```{r spagetti_appearance_each_hour}
plot_us_month = ufo_data %>% 
  #separate(date_time, into = c( "date","time"), sep = " " ) %>%
  separate( date, into = c("month","day","year"), sep = "/") %>% 
  filter(country == "USA",year >= 1950) %>%
  separate(time, into = c("hour", "minute"), sep = ":") %>% 
  mutate(hour = paste0(hour, ":00-", hour, ":59")) %>% 
  select(-minute) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  group_by(hour, ufo_shape) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = hour, y = n, color = ufo_shape)) +
  geom_line(aes(group = ufo_shape)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Number of UFO in Each Hour by Month in US",
    x = "Hour",
    y = "Number"
  ) +
  theme(plot.title = element_text(hjust = 0.5))

# display the plot
plot_us_month
```

*Conclusion:* 1):The most frequent shape of UFO is light; 
            2):The UFO is more likely to appear during night.


# Month/Year vs. Encounter length time 

```{r}
## USA
ufo =
ufo_data %>%  
separate( date, into = c("month","day","year"), sep = "/")

  ufo %>%
  filter(country == "USA", year >= 1950, encounter_length <6000) %>%
  select(year, month, encounter_length) %>%
  mutate(year = as.numeric(year),
         month = as.numeric(month)) %>%
  group_by(year,month) %>%
  arrange(desc(month)) %>%
  mutate(
  ave_encountered_length = mean(encounter_length/60)) %>%
  select(-encounter_length) %>%
  ungroup() %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  ggplot(aes(x = year, y = month)) +
  geom_tile(aes(fill = ave_encountered_length), colour = "white") +
  scale_fill_gradient(low = "lightyellow", high = "darkblue") +
  scale_x_continuous(name="Years", 
                     breaks = c(1950, 1955,1960,1965,1970,1975,1980,1985,1990,1995,2000,2005,2010,2015
                                ), limits= c (1950,2015)) +
  labs(
    title = "Average UFO Encounter Length Compared among Months Across Time in USA ",
    x = "Years",
    y = "Month"
  ) + 
  theme(axis.text.x = element_text(angle = 300)) +
  #theme(plot.subtitle = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(hjust = 0.5))
```

The above plot shows that there are more observations with lower encounter length time after the year of 1990. 

# Shape vs. Encounter length time 

<strong>The violin plot of encounter length for each shape of ufo & The bar chart of the median of encounter length time for each shape of UFO </strong>


```{r violin_plot_bar_chart}

# tidy data
length_shape_data = ufo_data %>% 
  mutate(ufo_shape = factor(ufo_shape),
         ufo_shape = forcats::fct_reorder(ufo_shape, encounter_length))

# violin plot 
length_shape_data %>% 
  ggplot(aes(x = ufo_shape, y = encounter_length, group = ufo_shape)) +
  geom_violin(aes(fill = ufo_shape), color = "cornsilk4", alpha = 0.4) +
  theme(axis.text.x = element_text(angle = 300)) +
  stat_summary(fun.y = median, geom = "point", color = "darksalmon", size = 3) +
  labs(title = "The Violin Plot of Encounter Length Time for Each Shape of UFO", 
       y = "Encounter Length of Time",
       x = "UFO Shape") +
  theme(plot.title = element_text(hjust = 0.5))

# tidy data for the bar chart
median_length_shape_data = ufo_data %>% 
  group_by(ufo_shape) %>% 
  summarize(median_length = median(encounter_length), 
            count = n()) %>% 
  filter(count > 100) %>% 
  arrange(desc(median_length)) %>% 
  ungroup()

# bar chart
median_length_shape_data %>% 
  ggplot(aes(x = ufo_shape, y = median_length, fill = ufo_shape)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 300)) +
  labs(title = "The Bar Chart of Median of Encounter Length Time for Each Shape of UFO",
       y = "Median of Encounter Length Time",
       x = "UFO Shape") +
  theme(plot.title = element_text(hjust = 0.5))

```

The violin plot here doesn't actually indicate much because the existence of outliers in several shapes, but we also can see that the medium (indicated by the pink dots) most of UFO encounter time length regardless of their shapes are small. 

The bar chart is a more in-depth view on median of UFO encounter length by shape. The median for the changing shape has the largest median (slightly over 400 seconds), and the smallest median of UFO encounter length is flash (slighter fewer than 50 seconds).


<strong>The Median Length Time for Each Shape of UFO (by day and night) </strong>

Whether the relationship between shape and the length time would change by time

```{r length_day_night}
ufo_day_night_length =
ufo_data %>%  
# separate(date_time, into = c( "date","time"), sep = " " ) %>%
separate(date, into = c("month","day","year"), sep = "/") %>%
separate(time, into = c("hour","minute"), sep = ":") %>% 
mutate(hour=recode(hour, "00"="24")) %>% 
filter(country == "USA") %>% 
mutate(year_1 = case_when(year >= 1950 ~ "true",
                          TRUE ~"false")) %>% 
filter(year_1 == "true") %>% 
select(-year_1) %>% 
mutate(hour = as.numeric(hour)) %>% 
mutate(night_day = case_when(18 <= hour & hour <= 24~ "night",
                             1 <= hour & hour <= 6 ~ "night",
                             TRUE ~ "day")) %>% 
  group_by(ufo_shape, night_day) %>% 
           summarize(median_length = median(encounter_length),
count = n()) %>% 
  filter(count > 100) %>% 
  ggplot(aes(x = ufo_shape, y = median_length, group = night_day)) +
  geom_col(aes(fill = night_day), alpha = .75, position = position_dodge(1)) +
  scale_fill_manual(
    values = c("lightblue","lightpink"), 
    labels = c("day","night")
    ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "The Median Length Time for Each Shape of UFO (by day and night)", 
       y = "median length",
       x = "UFO Shape") +
  theme(plot.title = element_text(hjust = 0.5))
  
  ufo_day_night_length
```

The plot did not display any particular pattern or trend over the shape of UFO, hence we conclude that there is no association between encounter length time and UFO shape. 


# UFO Appearance in Each Month
```{r ufo_appearance_month}
ufo_month_counts_plot =
ufo_data %>%  
# separate(date_time, into = c( "date","time"), sep = " " ) %>%
separate(date, into = c("month","day","year"), sep = "/") %>%
filter(country == "USA") %>% 
mutate(year_1 = case_when(year >= 1950 ~ "true",
                          TRUE ~"false")) %>% 
mutate(month = forcats::fct_relevel(month, c("1","2","3","4","5","6","7","8","9","10","11","12"))) %>% 
mutate(month = recode(month,"1"="Jan","2"="Feb","3"="Mar","4"="Apr","5"="May","6"="Jun","7"="Jul","8"="Aug","9"="Sep","10"="Oct","11"="Nov","12"="Dec")) %>% 
filter(year_1 == "true") %>% 
select(-year_1) %>% 
group_by(month) %>% 
summarize(count_month = n()) %>% 
  ggplot(aes(x = month, y = count_month)) +
  geom_col(aes(fill = month), alpha = .75, position = position_dodge(1)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "The Plot of the Appearance UFO for Each Month", 
       y = "Counts of Appearance",
       x = "Month") +
  theme(plot.title = element_text(hjust = 0.5))
  
ufo_month_counts_plot

```

According to the plot here, the number of appearance by month of UFO is lower during the first half of the year and higher during the second half of the year. 
In other words, the frequency of UFO appearance is higher after June, and peaks at July. 

*Conclusion:* 1):There is no association between the shape of UFO and encounter length time; 
            2):The UFO is more likely to appear during the second half of the year.
            

# *LASSO (length time vs. location, time, shape)*

After the visualization, we are interested in predicting and finding the association between the encounter length time with UFO shape, sighting location and sighting time. In order to make a accurate prediction, we use Lasso model to find the best fit. The reason for adapting Lasso model is that the parameter in the model will be large, and Lasso will give tune the parameters automatically into their best fit. 

```{r read dataset}
# Read Dataset
ufo =
  ufo_data %>%  
  filter(country == "USA") %>% 
  #na.omit(ufo_data) %>% 
  #separate(date_time, into = c( "date","time"), sep = " " ) %>%
  separate(date, into = c("month","day","year"), sep = "/") %>% 
  separate(time, into = c("hour","minute"), sep = ":") %>% 
  mutate(month = as.factor(month.name[as.numeric(month)]), year = as.numeric(year), hour = as.numeric(hour), minute = as.numeric(minute), day = as.numeric(day), country = as.factor(country), state = as.factor(state), city = as.factor(city)) %>% 
  mutate(add = paste(state, city)) %>% 
  select(year, month, day, hour, minute, state,  add, ufo_shape, encounter_length) %>% 
  na.omit() #%>% 
  

```
  
  
```{r define useful functions}
x_process = function(df){#get independent variables
  x = 
    df %>% 
    select(-encounter_length) %>% 
    model.matrix( ~ ., .)
  x
}



# make the response variables
y_process = function(df){#get dependent variables
  y = 
    df %>% 
    select(encounter_length) %>% 
    pull(encounter_length)
  y
}

```



```{r fig.width = 10, fig.height = 10}
# Run Lasso on the Data, Exclude City
ufo_state = 
  ufo %>% 
  select(-add)

lasso_fit_state = 
  glmnet::glmnet(x = x_process(ufo_state), y = y_process(ufo_state), alpha = 1) %>%  #alpha is the parameters to control the type of elastic net models, when alpha = 1, the elastic net model is equivalent to lasso.
  broom::tidy(.) # clean the result


# draw the plot
lasso_fit_state %>% 
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = lambda, estimate, color = term)) + 
  geom_line() + 
  scale_x_reverse() +
  labs(
    title = "Lineplot of the Lasso Estimated Coefficients for Different Lambda",
    x = "Lambda",
    y = "Estimated Coefficients"
    ) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 10, byrow = TRUE)) + 
  theme(plot.title = element_text(hjust = 0.5))

# cross - validation to calculate the best lambda
lambda_best_state = 
  glmnet::cv.glmnet(x = x_process(ufo_state), y = y_process(ufo_state), alpha = 1, nfolds = 10)


#best lambda
lambda_state = lambda_best_state$lambda.min

# fit the lasso model
lasso_best_fit_state = 
  glmnet::glmnet(x = x_process(ufo_state), y = y_process(ufo_state), alpha = 1, nlambda = 1, lambda = lambda_state) #alpha is the parameters to control the type of elastic net models, when alpha = 1, the elastic net model is equivalent to lasso.

# display the parameters
glmnet::coef.glmnet(lasso_best_fit_state) 

```

According to the result, none of terms are significant. Hence, we can conclude that the knowledge of state, shape, year, month, day, hour, minute does not contribute any insights on the encounter length of UFO.


## *Run lasso on data, only the city appeared more than 50 times included*

We only include cities that has UFO occured more than 50 times, because if not, we would need to adopt high-dimensional models which introduce non-significant terms in the model. To avoid insignificant model, we here only include cities with 50 or more UFO appearances.
```{r fig.width = 10, fig.height = 10}

ufo_city = 
  ufo %>% 
  group_by(add) %>% 
  mutate(count = n()) %>% 
  filter(count > 50) %>% 
  ungroup() %>% 
  select(-count) %>% 
  select(-state)
  
lasso_fit_city = 
  glmnet::glmnet(x = x_process(ufo_city), y = y_process(ufo_city), alpha = 1) %>%  #alpha is the parameters to control the type of elastic net models, when alpha = 1, the elastic net model is equivalent to lasso.
  broom::tidy(.) # clean the result


# draw the plot
lasso_fit_city %>% 
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = lambda, estimate, color = term)) + 
  geom_line() + 
  scale_x_reverse() +
  labs(
    title = "Lineplot of the Lasso Estimated Coefficients for Different Lambda",
    x = "Lambda",
    y = "Estimated Coefficients"
    ) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 12, byrow = TRUE)) + 
  theme(plot.title = element_text(hjust = 0.5))

# cross - validation to calculate the best lambda
lambda_best_city = 
  glmnet::cv.glmnet(x = x_process(ufo_city), y = y_process(ufo_city), alpha = 1, nfolds = 10)


#best lambda
lambda_city = lambda_best_city$lambda.min


# fit the lasso model
lasso_best_fit_city = 
  glmnet::glmnet(x = x_process(ufo_city), y = y_process(ufo_city), alpha = 1, nlambda = 1, lambda = lambda_city) #alpha is the parameters to control the type of elastic net models, when alpha = 1, the elastic net model is equivalent to lasso.

# display the parameters
glmnet::coef.glmnet(lasso_best_fit_city) 
```

According to the result, most of terms are not statistically significant. The only siginificant term is the city Virginia Beach. However, the coefficient for this term is too small (9.223037e-11), so that we can just ignore. Hence, we can conclude that the knowledge of state, city, shape, month, year, hour, day, mintue does not contribute any insights on the encounter length of UFO.


