---
title: "UFO_Prediction"
author: "Jungang Zou"
date: "11/29/2019"
output: html_document
---

```{r set up, include = FALSE}
# knitr will run the chunk but not include the chunk in the final document
# copy from Jeff

try(source("./dependencies.R"), silent = TRUE)

# ensure reproductivity
set.seed(10)

# load library

library(viridis)
library(ggridges)
library(patchwork)
library(rvest)
library(MLmetrics)
library(e1071)
library(tidyverse)
library(modelr)


knitr::opts_chunk$set(
  # display the code in the code truck above its results in the final document
  echo = TRUE,
  # do not display any warning messages generated by the code
  warning = FALSE,
  # set the figure to be 8 x 6, and the proportion it takes to be 90%
  fig.width = 8,
  fig.height = 6, 
  out.width = "90%"
)

# setting a global options for continuous data color family and a different format to set discrete data to have a color family
options(
  ggplot2.countinuous.colour = "viridis",
  ggplot2.countinuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# have a minimal theme and legends at the bottom
theme_set(theme_bw() + theme(legend.position = "bottom"))
```


```{r}
# read data
ufo_data = read_csv(file = "./data/tidied_data_final.csv")



# data tidy for prediction
ufo =
  ufo_data %>%  
  filter(country == "USA") %>% 
  separate(date, into = c("month","day","year"), sep = "/") %>% 
  separate(time, into = c("hour","minute"), sep = ":") %>% 
  select(year, month, day, hour, minute,state, city) %>%
  na.omit(ufo_data) %>% 
  mutate(add = paste(state, city)) %>% 
  mutate(month = as.factor(month.name[as.numeric(month)]), year = as.numeric(year), hour = as.numeric(hour), minute = as.numeric(minute), day = as.numeric(day))
```
  
  
```{r}  

# keep the city that the ufo appears with more than 100 times, and delete others
ufo_city = 
  ufo %>% 
  dplyr::group_by(add) %>% 
  mutate(count = dplyr::n()) %>% 
  filter(count >= 100) %>% 
  ungroup() %>% 
  select(-count) %>% 
  select(-state, -city) %>% 
  mutate(add = as.factor(add))

```


# Cross Validation Preparation

```{r}

# set up the result of cross validation table
nu_cv = tibble(
  nu = numeric(),
  f1_svm = list(),
  label = list(),
  pred_label = list()
)

# transform each levels in factors to new variables to form the input of SVM
factor_variable_cope = function(df){
  if (is.tibble(df)) {
    df = as.data.frame(df)
  }
  for (i in colnames(df[, sapply(df, is.factor)])) {
    for (level in unique(df[, i])) {
        df[paste(i, level, sep = "_")] = 
            as.integer(ifelse(df[, i] == level, 1, -1))
    }
  }
  df
}

```


# Cross Validation to tune the "nu" parameters for One Class Support Vector Machine

method for cross validation for one class data come from the paper:

Hempstalk, K., & Frank, E. (2008, December). Discriminating against new classes: One-class versus multi-class classification. In Australasian Joint Conference on Artificial Intelligence (pp. 325-336). Springer, Berlin, Heidelberg.


Use F1_score to evaluate the prediction result. F1_score can use to balance the result of imbalanced data.

https://en.wikipedia.org/wiki/F1_score


```{r}

# tune the parameter for nu in OC-SVM
for (nu_ in c(0.001, 0.005, 0.01, 0.05, 0.1, 0.3, 0.5, 0.7, 0.9, 0.99)) {
  false_samples = 5000 # numbers of false samples

  false_data = tibble( # false samples randomly generating
    year = sample(1900:2100, false_samples, replace = T),
    month = sample(ufo_city$month, false_samples, replace = T),
    day = sample(1:30, false_samples, replace = T),
    hour = sample(0:24, false_samples, replace = T),
    minute = sample(ufo_city$minute, false_samples, replace = T),
    add = sample(ufo_city$add, false_samples, replace = T)
  ) %>% 
    anti_join(ufo_city, by = c("year", "month", "day", "hour", "minute", "add")) # delete the samples of true in false dataset
  
  
  false_n = nrow(false_data) #numbers of false samples
  true_n = nrow(ufo_city) #numbers of true samples
  
  
  # here we bind rows of true and false dataset to deal with the factor variables
  train_df = 
    ufo_city %>% 
    bind_rows(false_data) %>% 
    factor_variable_cope() %>% 
    select(-month, -add) %>% 
    mutate(year = as.integer(year), day = as.integer(day), hour = as.integer(hour), minute = as.integer(minute)) %>% 
    sample_frac(1)
  
  
  # seperate the false and true samples from transformed data
  false_df = 
    train_df %>% 
    slice((true_n + 1):(true_n + false_n)) %>% 
    mutate(y = FALSE) %>% 
    as_tibble() 
  
  true_df = 
    train_df %>% 
    slice(1:true_n) %>% 
    mutate(y = TRUE) %>% 
    as_tibble() 
  
  
  # construct the cross validation dataset from true samples datasets
  cv_df = 
    crossv_mc(true_df, 10) %>% 
    mutate(
      train = map(train, as_tibble),
      test = map(test, as_tibble)
      )
  
  # add false samples to test dataset and construct the true labels 
  for (i in 1:nrow(cv_df)) {
    cv_df$test[[i]] = mutate(cv_df$test[[i]], y = TRUE)
    cv_df$test[[i]] = rbind(cv_df$test[[i]], false_df)
    cv_df$label[[i]] = cv_df$test[[i]]$y
    cv_df$test[[i]] = select(cv_df$test[[i]], -y)
  }
   
  # use OC-SVM to predict
  cv_model = 
    cv_df %>% 
    mutate(svm_mod = map(train, ~svm(.x , y = NULL,
             type = 'one-classification',
             nu = nu_, # tune the parameters by cross validation
             scale = TRUE, 
             kernel = "linear"))) %>% 
    mutate(pred_label = map2(svm_mod, test, ~predict(.x, .y)))
  
  # calculate the f1-score for each prediction
  cv_f1 =
    cv_model %>% 
    mutate(f1_svm = map2_dbl(.x = label, .y = pred_label, ~F1_Score(y_true = .x, y_pred = .y))) 
  
  # save the result 
  nu_cv = 
    tibble(
      nu = nu_,
      f1_svm = mean(pull(cv_f1, f1_svm))
    ) %>% 
    rbind(nu_cv, .)
  
}
```


draw the graph of f1_score for each nu

```{r}
# change the scale of data for graph
cvp = 
  nu_cv %>% 
  mutate(nu = case_when(
    nu <= 0.3 ~ log10(nu),
    nu > 0.3 ~ exp(nu)
  ))

# draw the graph
cvp %>% 
  ggplot(aes(x = nu, y = f1_svm, color = nu)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(
    breaks = pull(cvp, nu),
    labels = c("0.001", "0.005", "0.01", "0.05", "0.1", "0.3", "0.5", "0.7", "0.9", "0.99")
  ) +
  labs(
    title = "Cross Validation Plot for nu in SVM",
    x = "nu",
    y = "F1 Score") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```

From the graph we find nu = 0.5 is best parameter for OC-SVM.

After geting the best nu, we use it to train OC-SVM again.

```{r}
nu_ = 0.5

# train svm models
svm_model = 
    ufo_city %>% 
    bind_rows(false_data) %>% 
    factor_variable_cope() %>% 
    select(-month, -add) %>% 
    mutate(year = as.integer(year), day = as.integer(day), hour = as.integer(hour), minute = as.integer(minute)) %>% 
    sample_frac(1) %>% 
    svm(y = NULL,
             type = 'one-classification',
             nu = nu_,
             scale = TRUE, 
             kernel = "linear")
  
# display the models
svm_model

```





