---
title: "visualization"
author: "Keyi Wang"
date: "11/8/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(leaflet)
library(dplyr)
library(ggplot2)
library(ggmap)
library(sf)
library(mapview)
library(rgdal)
library(htmltools)


knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6, 
  out.width = "90%"
)


options(
  ggplot2.countinuous.colour = "viridis",
  ggplot2.countinuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

## reading data
```{r}
ufo_data = readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")

ufo =
ufo_data %>%  
na.omit(ufo_data) %>% 
separate(date_time, into = c( "date","time"), sep = " " ) %>%
separate( date, into = c("month","day","year"), sep = "/") %>% 
mutate(country = recode(country , 
                        "us" = "the United State", 
                        "au" = "Australia",
                        "gb" = "the United Kingdom",
                        "ca" = "Canada")) 
```


```{r}
## A bargraph showing the total number of UFO observed among 4 countries. As the number of UFO  observed in Australia and Britain is really low, for later graphs, we decided to omit data in these two countries.

ufo %>%
  group_by(country) %>%
  count() %>%
 mutate(precentage = n/66516) %>%
ggplot(aes(x = country, y = precentage , fill = country, label =  scales::percent(precentage))) + 
    geom_col(position = 'dodge') + 
    geom_text(position = position_dodge(width = .9),  
              vjust = -0.5,   size = 3) + 
    labs(title = "Bargraph of the precentage of UFO Observed by Country",
         x = "Country",y = "Precentage") 



```
Comment: 
According to the graph above, we can see that most of the data come from the United Statesa and some of them come from Canada. Australia and the United Kingdom contributed very litter amount of data. 

```{r}
## line plot showing the number of UFO observed across time among countries

ufo %>%
group_by(year) %>%
count(country) %>%
ungroup() %>%
mutate(year = as.numeric(year)) %>%
  ggplot(aes(x = year,y = n,group = country)) +
  geom_line(aes(color = country)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
title = " Line Plot of Number of UFO Observed Across Time Among different Countries ",
    x = "Year",
    y = "Number Of UFO Observed"
  )


```
Comment: 
In this plot, we can see that for the United States the earliest data happens before 1925 and sometime after 1960 the amount of UFO sighting begins to increase. We can also see that most of UFO sighting reports taken place around 1990 and after. Similar trend for Canada, but the amount of them is not as much as the United States. 

```{r}
### Boxplot showing the distribution of the UFO Shape Observed in US
  
ufo %>%
  filter(country == "the United State")%>%
  group_by(state) %>%
  count(ufo_shape) %>%
  arrange(desc(n)) %>% 
  ungroup %>% 
  mutate(
  ufo_shape = factor(ufo_shape),
  ufo_shape = forcats::fct_reorder(ufo_shape, n)) %>%
  ggplot(aes(x = ufo_shape, y = n, group = ufo_shape)) +
  geom_boxplot(aes(color =ufo_shape)) +
  theme(axis.text.x = element_text(angle = 300)) +
  labs(
    title = "Boxplot showing distribution of the UFO Shape Observed in US ",
    x = "UFO Shape",
    y = "Number"
  )
```
Comment: 
According to this plot, we can see that the light UFO shape was sighted most often, and the changed shape was the least often according to their medium value. Some of the most common sighted UFO shapes are light, trianle, circle, unknown, other, and fireball, disk, and sphere. Some of the least common sighted UFO shapes are changed crescent, delta, flare, hezagon, pyramid and round. 

```{r}
## heatmap showing 
  ufo %>%
  filter(year >= 1950, encounter_length < 10000) %>%
  select(year, month, encounter_length) %>%
  mutate(month = month.name[as.numeric(month)]) %>%
  group_by(year,month) %>%
  mutate(
  Ave_length_per_month = mean(encounter_length/60)) %>%
  select(-encounter_length) %>%
  ggplot(aes(x = year, y = month)) +
  geom_tile(aes(fill = Ave_length_per_month), colour = "white") +
  scale_fill_gradient(low = "lightyellow", high = "darkblue") +
  labs(
    title = "Average UFO Encounter Length Compared among Months Across Time ",
    x = "Years",
    y = "Month"
  ) + 
  theme(axis.text.x = element_text(angle = 90)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
  
```




```{r include=FALSE}

```


```{r}

  


```

